---
title: "results_for_meeting"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r read in data, include= FALSE}


setwd("C:/Users/obiew/Desktop/scott_pesticide_data")

# load packages
library(ggplot2)
library(reshape2)
library(gridExtra)
library(fitdistrplus)
library(lme4)
library(emmeans)
library(vegan)
library(dplyr)

rm(list = ls())

# DATA PREP -------------------
### read in data --------------------------------------------------

alldf<- read.csv("input/all_pesticide2019.csv")
ld50<- read.csv("input/ld50.csv")
pesticidenames<-read.csv("input/pesticide_type.csv")
gis <- read.csv("input/appleorchard_gis_data2015_3000m.csv")

#then fix some things
alldf$Sample.type <- as.factor(alldf$Sample.type)
pesticidenames$moa<-as.factor(pesticidenames$moa)
pesticidenames$pesticide<-as.factor(pesticidenames$pesticide)
pesticidenames$type<-as.factor(pesticidenames$type)
pesticidenames$class<-as.factor(pesticidenames$class)

# first lets only look at bees (may change this later)
alldf1 <- subset(alldf, Sample.type!="Apple flower" &
                           Sample.type!="Bumble bee wax"&
                           Sample.type!="Dandelion flower"&
                           Sample.type!="Nectar"&
                           Sample.type!="Honey bee pollen")


alldf1<-droplevels(alldf1) # drop unused levels
rownames(alldf1) = seq(length=nrow(alldf1)) # now fix row # after subsetting




### adjusting to HQ --------------------------------------------
# now we'll try to make everything in alldf (ppb) as a percent of ld50 of honey bees
# which would entitle dividing alldf by corresponding values of ld50

#drop identifier colummns
justpesticides <- alldf1[7:98]


ld50contact <- ld50[1,2:93]
ld50oral <- ld50[2,2:93]

# okay probably not the cleanest way but we can add the ld50 row to the pesticide df

# then matrix is and devide by the ld 50 row
oral<-rbind(justpesticides,ld50oral)

#row numbers got wonky again
rownames(oral) = seq(length=nrow(oral)) # now fix row # after subsetting


temp = as.matrix(oral)

oralhq<- t(t(temp)/temp[121,])

oralhq = as.data.frame(oralhq)

#now remove the ld 50 row
oralhq <- oralhq[1:120,]

# then we'll set all the inf and NA to 0 (where the ld50 was either 0 or NA)
oralhq[] <- lapply(oralhq, function(x) replace(x, !is.finite(x), 0))


#now do the same for contact
contact<-rbind(justpesticides,ld50contact)

#row numbers got wonky again
rownames(contact) = seq(length=nrow(contact)) # now fix row # after subsetting

temp = as.matrix(contact)

contacthq<- t(t(temp)/temp[121,])

contacthq = as.data.frame(contacthq)


contacthq <- contacthq[1:120,]

contacthq[] <- lapply(contacthq, function(x) replace(x, !is.finite(x), 0))



#now add back descriptors
oraldf<- cbind(alldf1[,1:6],oralhq)
contactdf<- cbind(alldf1[,1:6],contacthq)

#then on second thought lets drop some of those descriptor columns
oraldf<- cbind(alldf1[,c(3,4,6)],oralhq)
contactdf<- cbind(alldf1[,c(3,4,6)],contacthq)


# okay now just quickly double check the ug/g to ppb is correct
# first lets figure out the average weight of a hb in the sample
mean((oraldf$Mass..g.[oraldf$Sample.type == "Apis mellifera worker"]))

# okay .1 seems quite spot on





# now lets add a column to see total additive ld50 
#this assumes pesticides are purely additive
oraldf$chemsum <- rowSums(oraldf[4:95])
contactdf$chemsum <- rowSums(contactdf[4:95])


# and now add a column to see if any pesticides were found or not
oraldf$non_zero <- ifelse(oraldf$chemsum > 0, 1, 0)
contactdf$non_zero<- ifelse(contactdf$chemsum > 0, 1, 0)


#then quickly just check the frequency samples with no pesticides
table(oraldf$non_zero)
table(contactdf$non_zero)
# wow okay incredibly low



### remove undetected pesticides ------------------------------

# I think we should remove pesticides that were never detected
# or that have no ld50
# as theyre not informative in our ld50 data table
# and then just report which pesticides were removed in a table or something

# we can do this by summing columns and removing zeros 

i <- (colSums((oraldf[,4:95]), na.rm=T) != 0) # T if colSum is not 0, F otherwise
cols_to_drop_oral <- c(rep(TRUE, 3), i) # then add 3 T to keep meta data columns


oraldf1 <- oraldf[, cols_to_drop_oral] # all the non-zero columns plus meta data
notfoundpesticides.oral <- oraldf[, !cols_to_drop_oral]  # all the zero columns


# and the same for contact
i <- (colSums((contactdf[,4:95]), na.rm=T) != 0) # T if colSum is not 0, F otherwise
cols_to_drop_contact <- c(rep(TRUE, 3), i) # then add 3 T to keep meta data columns


contactdf1 <- contactdf[, cols_to_drop_contact] # all the non-zero columns plus meta data
notfoundpesticides.contact <- contactdf[, !cols_to_drop_contact]  # all the zero columns

# the below should sum to 97 if everything went well
# ncol(oraldf1) + ncol(notfoundpesticides.oral)
# ncol(contactdf1) + ncol(notfoundpesticides.contact)

# here is the list of pesticides that were not found or have no oral ld50
noimpactoral <- colnames(notfoundpesticides.oral)
noimpactcontact <- colnames(notfoundpesticides.contact)






### melting for graphing ----------------------------------
# now lets melt that data for graphing and playing with
oralmelt <- reshape2::melt(oraldf1[,1:34], id.vars=c("Mass..g.", "Sample.type", "Site"))

colnames(oralmelt)[4] <- "pesticide"
colnames(oralmelt)[5] <- "oralhq"


contactmelt <- reshape2::melt(contactdf1[1:41], id.vars=c("Mass..g.", "Sample.type", "Site"))

colnames(contactmelt)[4] <- "pesticide"
colnames(contactmelt)[5] <- "contacthq"


#now I would love to add pesticide type as a category

oralmelt <- merge(pesticidenames,oralmelt, by = "pesticide", all.y=TRUE)
contactmelt <- merge(pesticidenames,contactmelt, by = "pesticide", all.y=TRUE)

# then lets make a melt of only pesticide findings
oralmelt.findings <- subset(oralmelt, oralmelt$oralhq >0)
contactmelt.findings <- subset(contactmelt, contactmelt$contacthq >0)


### adding GIS information ####

# we have 2015 ag data for all 2019 orchards except 1 (chip bailey) 
gis_sum <- gis[,55:61]

gis_sum$Site <- as.factor(gis_sum$Site)

oralmelt$Site <- as.factor(oralmelt$Site)
contactmelt$Site <- as.factor(contactmelt$Site)
oraldf1$Site <- as.factor(oraldf1$Site)
contactdf1$Site <- as.factor(contactdf1$Site)


oralmelt <- merge(oralmelt,gis_sum, by = "Site", all.x = TRUE)
oraldf1 <- merge(oraldf1,gis_sum, by = "Site", all.x = TRUE)

contactmelt <- merge(contactmelt,gis_sum, by = "Site", all.x = TRUE)
contactdf1 <- merge(contactdf1,gis_sum, by = "Site", all.x = TRUE)




# okay lets clean things up
# and remove df I dont need

rm(list = c("alldf",
            "alldf1",
            "contact",
            "contacthq",
            "oralhq",
            "oral",
            "ld50",
            "ld50contact",
            "ld50oral",
            "notfoundpesticides.contact",
            "notfoundpesticides.oral",
            "i",
            "pesticidenames",
            "temp",
            "cols_to_drop_oral",
            "cols_to_drop_contact",
            "noimpactcontact",
            "noimpactoral",
            "justpesticides"
            ))





# END OF DATA PREP ####
```

# summary tables
```{r}
table(oraldf$Sample.type)
```
first the number of samples we have. this is about 1 per species per site

```{r}
table(oralmelt.findings$Sample.type,oralmelt.findings$type)
table(oralmelt.findings$Sample.type,oralmelt.findings$moa)

table(contactmelt.findings$Sample.type,contactmelt.findings$type)
table(contactmelt.findings$Sample.type,contactmelt.findings$moa)
```
we can see which pesticide classes are showing up in different bees in oral and contact hq


```{r}
table(oralmelt.findings$Sample.type)/table(oraldf$Sample.type)

```
and how many pesticides (oral) are found by sample type

```{r}
table(contactmelt.findings$Sample.type)/table(contactdf$Sample.type)
```
and for contact

```{r}

beecountvectororal <- c(19,20,16,16,20,17,12)

table(oralmelt.findings$Sample.type,oralmelt.findings$moa)/beecountvectororal

table(oralmelt.findings$Sample.type,oralmelt.findings$type)/beecountvectororal

```




```{r}
oraldf1 %>%
  group_by(Sample.type) %>% 
  summarise(mean_hq = mean(chemsum), max_hq = max(chemsum),perc_findings= mean(non_zero))


contactdf1 %>%
  group_by(Sample.type) %>% 
  summarise(mean_hq = mean(chemsum), max_hq = max(chemsum),perc_findings= mean(non_zero))



```
this is the mean values of hq (additive) and rate of findings




# now some overview summary graphs

```{r, echo=FALSE}
ggplot(oraldf1, aes(x=Sample.type, y=chemsum))+
  geom_boxplot(alpha=.5)+
  scale_y_continuous(trans='log10')+
  geom_hline(yintercept = 1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(oraldf1, aes(x=Sample.type, y=chemsum))+
  geom_boxplot(alpha=.5)+
  geom_hline(yintercept = 1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
here we see HB come out significantly higher than other sample types when looking at additive ORAL HQ

```{r, echo=FALSE}

ggplot(contactdf1, aes(x=Sample.type, y=chemsum))+
  geom_boxplot(alpha=.5)+
  scale_y_continuous(trans='log10')+
  geom_hline(yintercept = 1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(contactdf1, aes(x=Sample.type, y=chemsum))+
  geom_boxplot(alpha=.5)+
  geom_hline(yintercept = 1)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
when looking at additive contact HQ, HB and masonbee pollen both come out significantly higher than other sample types















```{r, echo=FALSE}
ggplot(oralmelt.findings, aes(x=Sample.type, y=oralhq, color=moa))+
  geom_boxplot(alpha=.5)+
  facet_wrap(~moa, scales="free_y")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(contactmelt.findings, aes(x=Sample.type, y=contacthq, color=moa))+
  geom_boxplot(alpha=.5)+
  facet_wrap(~moa, scales="free_y")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
here is the oral / contact hq by pesticide class (non additive zeros removed)



```{r, echo=FALSE}

ggplot(oralmelt, aes(x=perc_ag, y= oralhq, color=type))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~type)+
  scale_y_continuous(trans='log10')


ggplot(oralmelt, aes(x=perc_cornsoy, y= oralhq, color=type))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~type)+
  scale_y_continuous(trans='log10')

      
ggplot(oralmelt, aes(x=perc_apple, y= oralhq, color=type))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~type)+
  scale_y_continuous(trans='log10')



ggplot(oraldf1, aes(x=perc_ag, y= chemsum, color=Sample.type))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Sample.type)+
  facet_wrap(~Sample.type, scales = "free_y")+
  scale_y_continuous(trans='log10')

ggplot(oraldf1, aes(x=perc_cornsoy, y= chemsum, color=Sample.type))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Sample.type)+
  facet_wrap(~Sample.type, scales = "free_y")+
  scale_y_continuous(trans='log10')



ggplot(oraldf1, aes(x=perc_apple, y= chemsum, color=Sample.type))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Sample.type, scales = "free_y")+
  scale_y_continuous(trans='log10')




ggplot(oraldf1, aes(x=perc_ag, y= chemsum))+
  geom_point()+
  geom_smooth()




```









